---
title: "quiz4"
author: "林振炜"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Chapter 4





## Example 1  Horseshoe crab grouped data logistic regression with Width
```{r warning=FALSE, paged.print=FALSE}
library(cdabookdb)
library(cdabookfunc)
library(MASS)
library(dplyr)
data("horseshoecrabs")
```


### Question 1: Fit logistic regression with ungrouped data with Width only.
```{r}
horseshoecrabs$psat <- as.integer(horseshoecrabs$Satellites > 0)
logit_model <- glm(psat~Width,data = horseshoecrabs,family = binomial(link = 'logit'));summary(logit_model)
```


### Question 2: Fit logistic regression with grouped data with width only.
```{r warning=FALSE, paged.print=FALSE}
horseshoecrabs$width_group <- cut(horseshoecrabs$Width,
                                  breaks = c(0, 23.25 + 0:6, Inf),
                                  dig.lab = 4)

mean_width_vs_prop <- sapply(levels(horseshoecrabs$width_group),function(x){
  # Declare that is grouped by width_group
  sub <- subset(horseshoecrabs,width_group==x)
  c(mean(sub$psat),mean(sub$Width))
})
prop <- mean_width_vs_prop[1,]
mean_width <- mean_width_vs_prop[2,]
logit_model <- glm(prop~mean_width,family = binomial(link = 'logit'));summary(logit_model)
```

### Question 3: Repeat Question 1 and 2 on Weight only.
```{r,warning=F}
summary(horseshoecrabs$Weight)
logit_model <- glm(psat~Weight,data = horseshoecrabs,family = binomial(link = 'logit'));summary(logit_model)


horseshoecrabs$Weight_group <- cut(horseshoecrabs$Weight,
                                  breaks = c(0, 1.2 + 0:4, Inf),
                                  dig.lab = 4)#dig.lab是一个用于确认分割线精度的参数


mean_Weight_vs_prop <- sapply(levels(horseshoecrabs$Weight_group),function(x){
  # Declare that is grouped by Weight_group
  sub <- subset(horseshoecrabs,Weight_group==x)
  c(mean(sub$psat),mean(sub$Weight))
})
prop <- mean_Weight_vs_prop[1,]
mean_Weight <- mean_Weight_vs_prop[2,]
logit_model <- glm(prop~mean_Weight,family = binomial(link = 'logit'));summary(logit_model)
```

### Question 4: Fit logistic regression with ungrouped data. 
              # (1) width and color, color is dummy/indicator.
              # (2) width and color, color is score.
              # (3) width and color, color is binary(dark or not).


```{r}
horseshoecrabs <- horseshoecrabs %>%
  mutate(
    Color_factor = factor(Color, 4:1) # Convert Color to a factor and set the factor level
    )
m1 <- glm(psat ~ Width + Color_factor, data = horseshoecrabs, family = binomial());summary(m1)

m2 <- glm(psat~Width+Color,data = horseshoecrabs,family = binomial());summary(m2)
horseshoecrabs$is_dark <- as.character(horseshoecrabs$Color < 4)
m3 <- glm(psat ~ is_dark + Width,
          family = binomial(),
          data = horseshoecrabs)
summary(m3)
```

### Question 5: with interaction
              # (1) width and color, color is dummy/indicator.
              # (2) width and color, color is score.
              # (3) width and color, color is dark or not.
```{r}
m1 <- glm(psat ~ Width + Color_factor+Width*Color_factor, data = horseshoecrabs, family = binomial());summary(m1)

m2 <- glm(psat~Width+Color+Width*Color,data = horseshoecrabs,family = binomial());summary(m2)
horseshoecrabs$is_dark <- as.character(horseshoecrabs$Color < 4)
m3 <- glm(psat ~ is_dark + Width+is_dark * Width,
          family = binomial(),
          data = horseshoecrabs)
summary(m3)
```




##  Example 2: AZT  logsitc with categorical predictors


### Question 1: fit logistic regression with Y and Treat and Race, no interactoin.
```{r}
library(tidyr)
data("AZT")
AZT_df <- spread(as.data.frame(AZT), Symptoms, Freq)
AZT_df
m <- glm(
cbind(Yes, No) ~ (Race == "White") + (AZTUse == "Yes"),
data = AZT_df,
family = binomial("logit")
)
summary(m)
```


### Question 2: Use three-way table, analysis the conditional independence between Y and Treatment for two levels of Race, separately.

```{r}
white<-AZT[1,1:2,1:2]
independent_test_of_table(white, "X2")
black<-AZT[2,1:2,1:2]
independent_test_of_table(black,'X2')

```

### Question 3: Conduct CMH test for AZT data.

```{r}
library(cdabookfunc)
cmh.test(cond="Race",data=AZT)
```





