---
title: "Practice_Chapter6"
author: "Becky"
date: "12/10/2019"
output: html_document
---
```{r  Chapter 6}
```

# Example 1: Alligator Example by Baseline-category logit models

# Question 1: Building baseline-category logit model with Food ~ Length

```{r}
library(VGAM)
library(nnet)
library(VGAMdata)
library(icda)

data(alligators1) 

attach(alligators1)

# baseline is "F"

fitm1<- vglm(Food ~ Length,family=multinomial, data=alligators1)
summary(fitm1) 

# baseline is "O"
fFood<-factor(Food,level=c("O","I","F"))

alligators1$fFood <- fFood

fitm2<- vglm(fFood ~ Length, family=multinomial, data=alligators1)
summary(fitm2)
```

# Question 2: Baseline is "I"
```{r}
fFood<-factor(Food,level=c("I","F","O"))

alligators1$fFood_I <- fFood

fitm3<- vglm(fFood_I ~ Length, family=multinomial, data=alligators1)
summary(fitm3)

fitm4<- vglm(fFood_I ~ Length, family=multinomial, data=alligators1)
summary(fitm4)
```


# Example 2: Belief in Afterlife Example by Baseline-category logit models Table 6.4
# Question 1: Fit baseline-category logit model  Believe ~ Gender + Race 

```{r}
data(afterlife) 

attach(afterlife)

fitm5<- vglm(Believe ~ Gender+Race,weights=Freq,family = multinomial, data=afterlife)
summary(fitm5)

```

# Question 2: "No" is baseline, and Female = 1, White =1 

```{r}
# "No" is baseline
fBelieve<- factor(Believe, level=c("No","Yes","Undecided"))

# Female =1
fGender <- factor(Gender, level=c("Male","Female"))

# White =1 
fRace <- factor(Race, level=c("Black","White"))

afterlife$fRace <- fRace
afterlife$fGender <- fGender
afterlife$fBelieve <- fBelieve

fitm6<- vglm(fBelieve ~ fGender+fRace,family= multinomial, weights=Freq,data=afterlife)
summary(fitm6)
fitted(fitm6)
```

# Example 3: Political Ideology and Party Affiliation Example 
# Question 1: Fit this data by Cumulative Logit Models with Proportional Odds Property

```{r}
data(ideology) 

attach(ideology)

ideology

#proportional odds model

orderIdeology <- ordered(Ideology,  levels = c("VCon","SCon","Mod","SLib","VLib"))

fit.order <- vglm(orderIdeology~ Party + Gender,weights=Freq,
                  family=cumulative(parallel=TRUE), data=ideology)
summary(fit.order)

```

# Question 2: Male =1, Rep = 1

```{r}
fit.order2 <- vglm(orderIdeology~ (Party=="Rep") + (Gender=="Male"),weights=Freq,
                  family=cumulative(parallel=TRUE), data=ideology)
summary(fit.order2)
```

# Example 4: Mental Health Data
# Question 1: Fit this data by Cumulative logit model with ungrouped data with proportional odds

```{r}
library(cdabookdb)
data("impairment")
attach(impairment)
fit7 <- vglm(Impairment ~ SES + LifeEvents,
            family=cumulative(parallel=TRUE), data=impairment)
summary(fit7)
```

# Question 2: Fit this data by Cumulative logit model with ungrouped data with non-proportional odds
```{r}
fit8 <- vglm(Impairment ~ SES + LifeEvents,
             family=cumulative, data=impairment)
summary(fit8)
```

# Question 3: Compare models in Question 1 and 2, which one is better, why?
```{r}
# residual deviance fit7  99.0979 on 115 
# residual deviance fit8  96.7486 on 111

pchisq(deviance(fit7)-deviance(fit8),df=df.residual(fit7)-df.residual(fit8),lower.tail=FALSE)
# can not reject H0, so fit7 is better.

```

# Example 5: Ideology by  Adjacent-Categories Logits
# Question 1: Fit the data by Adjacent-Categories Logits with proportional odds

```{r}
data(ideology) 

ideology

#proportional odds model

orderIdeology <- ordered(Ideology,  levels = c("VCon","SCon","Mod","SLib","VLib"))

# Female =1
fGender <- factor(Gender, level=c("Male","Female"))

# Dem =1 
fParty <- factor(Party, level=c("Rep","Dem"))

fitideo1.acat <- vglm(orderIdeology~ fParty + fGender,weights=Freq,
                      family=acat(reverse=TRUE, parallel=TRUE), data=ideology)
summary(fitideo1.acat)
```

# Question 2: Fit the data by Adjacent-Categories Logits with non-proportional odds
```{r}
fitideo2.acat <- vglm(orderIdeology~ fParty + fGender,weights=Freq,
                      family=acat(reverse=TRUE), data=ideology)
summary(fitideo2.acat)
```


# Example 6: Mental Health by  Continuation-Ratio Logits
# Question 1: Fit this data by Continuation-Ratio Logits
```{r}
data("impairment")
attach(impairment)
orderImpairment<- ordered(Impairment,  levels = c("Well","Mild","Moderate","Impaired"))

fit9.cratio <- vglm(orderImpairment ~ SES + LifeEvents,
                    family=cratio(reverse=FALSE, parallel=TRUE), data=impairment)
summary(fit9.cratio)

```

# Example 7: Jobsatisfaction Example
# Question 1: Build Cumulative logit model with Job~ Gender + Income

```{r}
data(jobsatisfaction) 

attach(jobsatisfaction)

jobsatisfaction

orderJob <- ordered(JobSat,  levels = c(1,3,4,5))

fitJob.order <- vglm(orderJob~ Gender + Income,weights=(Freq+0.1),
                     family=cumulative(parallel=TRUE), data=jobsatisfaction)

summary(fitJob.order)
```

# Question 2: Analysis the effect of Income according to the output

```{r}
abs(coef(fitJob.order)[5]/0.018)>1.96
# reject H0, so Income has effect given gender

```

# Question 3: Build Baseline category logit model with Job~ Gender + Income
```{r}
fitJob.cum <- vglm(JobSat~ Gender + Income,weights=(Freq+0.1),
                     family=multinomial, data=jobsatisfaction)

summary(fitJob.cum)
```

# Question 4: Compare models in question 1 and 3, which one is better, why?

```{r}
pchisq(deviance(fitJob.order)-deviance(fitJob.cum),df=df.residual(fitJob.order)-df.residual(fitJob.cum),lower.tail=FALSE)

# can not reject H0. So, cumulative is better.
```



 